---
title: "Module 3"
author: "Tyler"
date: "2023-03-03"
output: pdf_document
---
```{r}
library(tidyverse)
library(tibble)
library(reshape2)
library(stringr)
library(plotly)
library(shiny)
library(rsconnect)
```

```{r}
df = read.csv('https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv')
head(df)
```

### Question 1

As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.

```{r}
df_2010 = df %>%
  filter(., Year == "2010" & ICD.Chapter == "Endocrine, nutritional and metabolic diseases")
head(df_2010)
```

```{r}
enmd = df_2010 %>%
  plot_ly(x = ~df_2010$Crude.Rate, y = df_2010$State, type = "bar")
subplot(enmd)
```

```{r}
df_2010 = subset(df, Year==2010)
cods = unique(df_2010$ICD.Chapter)

shiny_1 = fluidPage(
  titlePanel("2010 Mortality Rates"),
  fluidRow(selectInput("cod", "Cause of Death:", choices=sort(cods))), plotOutput("plot")
)

cods = as.data.frame(unique(df_2010$ICD.Chapter))

server_1 = function(input,output) {
  rates = reactive({rates = subset(df_2010, ICD.Chapter==input$cod)})
  output$plot = renderPlot({
    ggplot(rates(), aes(x=Crude.Rate, y=reorder(State, -Crude.Rate)))+
      scale_x_continuous(limits=c(0,max(rates()$Crude.Rate))+5, expand=c(0,0))+
      geom_segment(aes(yend=State),xend=0,color="red")+
      geom_point(size=3,color="black")+
      theme_bw()+
      theme(panel.grid.major.y = element_blank(), axis.title=element_text(size=14))+
      xlab("2010 Mortality Rate")+
      ylab("State")
  })
}
```

```{r}
shinyApp(ui = shiny_1, server= server_1)
```

### Question 2

Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at a time. Keep in mind that the national average should be weighted by the national population. 

```{r}
shiny_2 = fluidPage(
  titlePanel("State Mortality Rates"),
  sidebarPanel(
    selectInput('cod2', 'Cause of Death:', 
                choices=sort(cods2), selected=1),
    selectInput('states', 'State', 
                choices=sort(states), selected=1)
  ),
  mainPanel(
    plotOutput('plot3')
  ))

server_2 = function(input, output) {
  output$plot3 = renderPlot({
  df %>%
      group_by(Year, ICD.Chapter) %>%
      mutate(tot_pop = sum(Population),
             tot_count = sum(Deaths),
             tot_crude_rate = 100000*(tot_count/tot_pop)) %>%
      group_by(Year, ICD.Chapter, State) %>%
      mutate(sum_count = sum(Deaths),
             sum_crude_rate = 100000*(sum_count/Population)) %>%
      select(ICD.Chapter, State, Year, tot_crude_rate, sum_crude_rate) %>%
      filter(ICD.Chapter == input$cod2, State == input$states) %>%
      ggplot() +
      geom_bar(aes(x=Year, weight = sum_crude_rate)) +
      labs(x='State', y='Mortality Rate')+
      geom_line(aes(x=Year, y=tot_crude_rate, linetype="National Average"), col="blue")
  })
}
```

```{r}
shinyApp(ui = shiny_2, server = server_2)
```